---
format:
  pdf:
    documentclass: article
    pdf-engine: xelatex  # Required for system fonts
    geometry:
      - top=0.5in
      - bottom=0.5in
      - left=0.5in
      - right=0.5in
    fontsize: 11pt
    colorlinks: true
    fig-pos: 'H'
   
    include-in-header:
      text: |
        \newcommand{\districtname}{Charlottesville}
        \usepackage{multicol}
        \usepackage{paracol}
        \usepackage{xcolor}
        \usepackage{tcolorbox}
        \usepackage{graphicx}
        \usepackage{float}
        \definecolor{maroon}{HTML}{601215}
        \usepackage{fontspec}
        \newfontfamily\leaguegothic{League Gothic}
        \newfontfamily\lato{Lato}
        \pagestyle{empty}
    mainfont: "Lato"  # Main body text font
execute:
  echo: false
  warning: false
  message: false

---
```{r setup}
#| include: false
library(knitr)
library(ggplot2)
library(dplyr)
library(here)

# Set figure defaults
knitr::opts_chunk$set(
  fig.width = 6,
  fig.height = 4,
  dpi = 300,
  out.width = "100%"
)



data_path <- "/Users/jfischman/Library/CloudStorage/Dropbox-TheLab/Josie Fischman/Github/gridded_eif_data/"
git_path <- "/Users/jfischman/Library/CloudStorage/Dropbox-TheLab/Josie Fischman/Github/environmental-inequality-atlas"

# Load packages

if (!requireNamespace("pacman", quietly = TRUE)) {
  install.packages("pacman")
}

pacman::p_load(ncdf4, sf, terra, dplyr, ggplot2, tmap, arrow, dplyr, tidyr, scales, haven, stringr, purrr, tigris, here, showtext, patchwork, cowplot)


raceincome_path <- here::here("data", "processed", "raceincome_rda", "nat_raceincome_2023.rda")
raceincome_2023 <- load(raceincome_path)
raceincome2023 <- get(raceincome_2023)

# Compute summaries
race_summary2023 <- raceincome2023 %>%
  filter(
    !is.na(race_ethnicity),
    !is.na(income_decile)
  ) %>%
  group_by(race_ethnicity, income_decile) %>%
  summarise(
    total_count = sum(n_noise_postprocessed, na.rm = TRUE),
    .groups = "drop"
  )

# U.S. Summary
us_summary <- raceincome2023 %>%
  filter(!is.na(race_ethnicity), !is.na(income_decile), !race_ethnicity %in% c(NA, "NA", "Unknown", "Other/Unknown"), income_decile != 0) %>%
  group_by(region = "United States", race = race_ethnicity, decile = income_decile) %>%
  summarise(total_count = sum(n_noise_postprocessed, na.rm = TRUE), .groups = "drop")


# Charlottesville/Albemarle Summary
# Load ONLY District 5 data from 2024 file (if file is big)
raceincome_path_2024 <- here::here("data", "processed", "raceincome_rda", "raceincome_2024.rda")

# Load into a temp environment so you don’t load everything
temp_env <- new.env()
load(raceincome_path_2024, envir = temp_env)

# Find the object name in that file
obj_name_2024 <- ls(temp_env)
raceincome2024 <- temp_env[[obj_name_2024]]


# Now filter for District 5 only
district_summary <- raceincome2023 %>%
  filter(
    STATEFP == "51",
    COUNTYFP %in% c("003", "540"), 
    !is.na(race_ethnicity),
    !is.na(income_decile),
    !race_ethnicity %in% c("NA", "Unknown", "Other/Unknown"),
    income_decile != 0
  ) %>%
  group_by(region = "Charlottesville", race = race_ethnicity, decile = income_decile) %>%
  summarise(
    total_count = sum(n_noise_postprocessed, na.rm = TRUE),
    .groups = "drop"
  )

# Combine and process
combined_summary <- bind_rows(us_summary, district_summary)

combined_prop <- bind_rows(us_summary, district_summary) %>%
  mutate(region = as.character(region)) %>%
  group_by(region, race) %>%
  mutate(share = total_count / sum(total_count)) %>%
  ungroup()

combined_prop$race <- factor(combined_prop$race, levels = c("White", "Black", "Hispanic", "Asian", "AIAN"))

us_data <- combined_prop %>% filter(region == "United States") %>% mutate(alpha = 0.5)
district_data <- combined_prop %>% filter(region == "Charlottesville") %>% mutate(alpha = 1)

avg_decile_by_race <- combined_summary %>%
  group_by(region, race) %>%
  summarise(avg_decile = weighted.mean(decile, total_count, na.rm = TRUE), .groups = "drop")

avg_decile <- combined_summary %>%
  group_by(region) %>%
  summarise(avg_decile = weighted.mean(decile, total_count, na.rm = TRUE), .groups = "drop")

district_decile_summary <- raceincome2024 %>%
  filter(
    STATEFP == "51",
    COUNTYFP %in% c("003", "540"),,
    !is.na(income_decile),
    income_decile != 0,
    !is.na(n_noise_postprocessed)
  ) %>%
  mutate(region = "Charlottesville") %>%
  group_by(region, decile = income_decile) %>%
  summarise(
    total_count = sum(n_noise_postprocessed, na.rm = TRUE),
    .groups = "drop"
  )

# From the 2023 dataset, get national summary
us_decile_summary <- raceincome2023 %>%
  filter(
    !is.na(income_decile),
    income_decile != 0,
    !is.na(n_noise_postprocessed)
  ) %>%
  mutate(region = "United States") %>%
  group_by(region, decile = income_decile) %>%
  summarise(
    total_count = sum(n_noise_postprocessed, na.rm = TRUE),
    .groups = "drop"
  )

# Combine U.S. and District 5
decile_summary <- bind_rows(us_decile_summary, district_decile_summary)

# Calculate share
decile_prop <- decile_summary %>%
  group_by(region) %>%
  mutate(share = total_count / sum(total_count)) %>%
  ungroup()

font_add(family = "Lato", regular = "/Users/jfischman/Library/CloudStorage/Dropbox-TheLab/Josie Fischman/Github/environmental-inequality-atlas/code/quarto/Lato,League_Gothic (1)/Lato/Lato-Regular.ttf")  # optional, if not auto-detected
showtext_auto()


```


```{=latex}


{\leaguegothic\textcolor{maroon}{{\Huge Income}}}

```




```{=latex}


{\lato\textcolor{maroon}{\textbf{\large \districtname~and Albemarle County Trends and Shares}}}

```
30\% of the \districtname{} population is in the top 2 deciles. The average income in 2024 was \$197,636. The average annual growth is 1.4\%.


```{=latex}



{\lato\textcolor{black}{\textbf{\large \districtname~and Albemarle County Income Decile Distribution}}}
\vspace{6pt}
```

```{=latex}
\setlength{\columnsep}{20pt}
\columnratio{0.6}
\begin{paracol}{2}
```

<!-- LEFT COLUMN -->

```{r income}

#| fig-cap: "Income Decile Distribution in Charlottesville"
#| fig-width: 9
#| fig-height: 4.5
#| warning: false
#| message: false




ggplot(decile_prop %>% filter(region == "Charlottesville"),
             aes(x = factor(decile), y = share)) +
  
  geom_col(
    fill = "#003f5c",
    width = 0.9
  ) +
  geom_hline(yintercept = 0, color = "black", size = 0.5) +
  # Labels and themes
  labs(
    title = "",
    x = "Income Decile",
    y = "Share of Total Population"
  ) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  theme_minimal(base_family = "Lato") +
  theme(
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
    legend.position = "none",
    panel.grid.major.x = element_blank(),  # removes vertical gridlines
    panel.grid.minor.x = element_blank(),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    axis.text = element_text(size = 12),
    
  )

```

```{=latex}
\begin{tcolorbox}[
  colback=white,
  colframe=white,
  boxrule=1pt,
  arc=0pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt,
  width=\linewidth
]
\color{black}
\scriptsize
\begin{minipage}{\dimexpr\linewidth-16pt}
\textit{Notes:} The figure shows the share of the \districtname~population within each decile of the national income distribution.
  
\end{minipage}


\end{tcolorbox}


```



```{=latex}
\switchcolumn
```

<!-- RIGHT COLUMN -->

```{=latex}




\begin{tcolorbox}[
  colback=maroon,
  colframe=maroon,
  boxrule=1pt,
  arc=0pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt,
  width=\linewidth
]
\color{white}
\begin{center}
\textbf{Income Deciles:}\\[12pt]

Income deciles divide the national population into ten equal income bins. The first decile reflects the 10\% of the U.S. population with the lowest income, while the 10th decile reflects the 10\% of the population with the highest income. The figure shows how the \districtname~income distribution differs from the U.S. A positive slope indicates a higher share of residents in the higher deciles while a negative slope indicates a higher share of residents in the lower deciles.  


\end{center}
\end{tcolorbox}





```

```{=latex}
\end{paracol}

```

```{=latex}

{\lato\textcolor{black}{\textbf{\large \districtname~Income Table}}}

\vspace{-0.5em}
```

\includegraphics[width=1\linewidth, trim=0.5cm 2cm 2cm 0.5cm, clip]{/Users/jfischman/Library/CloudStorage/Dropbox-TheLab/Josie Fischman/Github/environmental-inequality-atlas/output/raceincome/income_table.png}


```{=latex}
\vspace{-1em}
\begin{tcolorbox}[
  colback=white,
  colframe=white,
  boxrule=1pt,
  width=\textwidth,  % full width
  arc=0pt,
  left=6pt, right=6pt, top=4pt, bottom=4pt
]
\color{black}
\scriptsize
\begin{minipage}{\dimexpr\linewidth-16pt}
\textit{Notes:} The table shows the average income level for each decile group in 2024.The annual growth represents the average annual change in income from 1999 to 2024. Positive values represent an increase in average income while negative values represent a decrease in average income.
\end{minipage}
\end{tcolorbox}
```

```{=latex}
\newpage


{\leaguegothic\textcolor{maroon}{{\Huge Race and Income}}}


```

```{=latex}
\vspace{1pt}


{\lato\textcolor{maroon}{\textbf{\large \districtname~Race-Income Trends and Shares}}}

```
The population with the highest average income is the White population and the population with the lowest average income is the Black population. The population with the highest average annual income growth rate from 1999-2024 is the Asian population and the population with the lowest average annual growth rate is the Black population.
```{=latex}


{\lato\textcolor{black}{\textbf{\large \districtname~Race-Income Table}}}

\vspace{-0.5em}
```

\includegraphics[width=1\linewidth, trim=0.5cm 2cm 2cm 0.5cm, clip]{/Users/jfischman/Library/CloudStorage/Dropbox-TheLab/Josie Fischman/Github/environmental-inequality-atlas/output/raceincome/cville_growth_table.png}


```{=latex}
\vspace{-1em}
\begin{tcolorbox}[
  colback=white,
  colframe=white,
  boxrule=1pt,
  width=\textwidth,  % full width
  arc=0pt,
  left=6pt, right=6pt, top=6pt, bottom=6pt
]
\color{black}
\scriptsize
\begin{minipage}{\dimexpr\linewidth-16pt}
\textit{Notes:} The table shows that average income and the average growth rate over the last 25 years by race. Higher growth rates indicate faster growth over time. 

\end{minipage}
\end{tcolorbox}
```


```{=latex}


{\lato\textcolor{black}{\textbf{\large Race Specific Income Distributions: \districtname~and Albemarle vs. U.S.}}}

```

```{r income by race}
#| warning: false
#| message: false
#| fig-width: 9
#| fig-height: 4.5

race_order <- c("White", "Black", "Hispanic", "Asian", "AIAN")

library(ggplot2)
library(patchwork)
library(grid)        # for grid.draw
library(cowplot)     # for get_legend

# Your plotting function without legend:
plot_income_by_race <- function(df, race_name, show_legend = FALSE) {
  df_filtered <- df %>% filter(race == race_name)
  
  p <- ggplot(df_filtered, aes(x = as.numeric(decile), y = share, fill = region)) +
    geom_col(position = position_dodge2(padding = 0), width = 0.7) +
    geom_hline(yintercept = 0, color = "black", size = 0.5) +
    scale_x_continuous(breaks = 1:10, labels = as.character(1:10)) +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
    scale_fill_manual(values = c("United States" = "#a6cee3", "Charlottesville" = "#003f5c")) +
    labs(
      title = race_name,
      x = "Income Decile",
      y = "Share of Race Group",
      fill = "Region"
    ) +
    theme_minimal(base_size = 14, base_family = "Lato") +
    theme(
      plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
      axis.title.x = element_text(size = 12),
      axis.title.y = element_text(size = 8),
      axis.text = element_text(size = 8),
      panel.background = element_blank(),
      plot.background = element_rect(fill = "white", color = NA),
      panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank(),
      panel.grid.major.y = element_line(color = "gray90"),
      legend.position = if (show_legend) "right" else "none"   # legend only on last plot, on right side
    )
  
  return(p)
}

# Create plots:
plots_no_legend <- lapply(race_order[-length(race_order)], function(r) plot_income_by_race(combined_prop, r, FALSE))
last_plot_with_legend <- plot_income_by_race(combined_prop, race_order[length(race_order)], TRUE)

# 2. Extract legend from last plot
legend <- get_legend(last_plot_with_legend)

# 3. Remove legend from last plot
last_plot_no_legend <- last_plot_with_legend + theme(legend.position = "none")

# 4. Combine all 5 plots (no legends)
all_plots <- c(plots_no_legend, list(last_plot_no_legend))

# 5. Create an "empty" blank plot for the 6th cell where the legend will go
empty_plot <- ggplot() + theme_void()

# 6. Build the grid with 6 plots (5 plots + 1 empty)
plots_grid <- plot_grid(plotlist = c(all_plots, list(empty_plot)),
                        ncol = 3,
                        align = "hv")

# 7. Overlay the legend **on top of the empty plot** cell using cowplot’s `ggdraw()`

# Create a drawing canvas over the whole grid
final_with_legend <- ggdraw(plots_grid) +
  draw_grob(legend, 
            x = 0.67,  # roughly center in the last column (3rd column)
            y = 0.05,        # center in the bottom row (2nd row)
            width = 1/3,    # width covering about one cell width
            height = 1/2)   # height covering about one cell height

print(final_with_legend)
```

```{=latex}
\begin{tcolorbox}[
  colback=white,
  colframe=white,
  boxrule=1pt,
  width=\textwidth,  % full width
  arc=0pt,
  left=6pt, right=6pt, top=4pt, bottom=4pt
]
\color{black}
\scriptsize
\begin{minipage}{\dimexpr\linewidth-16pt}
\textit{Notes:} The dark blue bars show the share of the population within each decile of the national income distribution for each race in \districtname~while the light blue bars show the income decile distributions for every race in the U.S. The set of figures shows how race specific income distributions in \districtname~differ from U.S. race specific income distributions. A positive slope indicates a higher proportion of individuals in the upper deciles while a negative slope indicates a higher proportion of individuals in the lower deciles. Differences between the dark and light blue lines indicate a difference in income distribution for the given decile.

\end{minipage}
\end{tcolorbox}
```


  



