---
format:
  pdf:
    documentclass: article
    pdf-engine: xelatex  # Required for system fonts
    geometry:
      - top=0.5in
      - bottom=0.5in
      - left=0.5in
      - right=0.5in
    fontsize: 11pt
    colorlinks: true
    fig-pos: 'H'
    mainfont: "Lato"  # Main body text font
    include-in-header:
      text: |
        \newcommand{\districtname}{District 5}
        \usepackage{multicol}
        \usepackage{paracol}
        \usepackage{xcolor}
        \usepackage{tcolorbox}
        \usepackage{graphicx}
        \usepackage{float}
        \definecolor{maroon}{RGB}{128,0,0}
        \usepackage{fontspec}
        \usepackage{datetime}
        \newfontfamily\leaguegothic{League Gothic}
        \newfontfamily\lato{Lato}
        % Suppress page numbers
        \pagestyle{empty}
        \definecolor{custommaroon}{HTML}{601215}
        \usepackage[none]{hyphenat}
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
#| include: false
library(knitr)
library(ggplot2)
library(dplyr)

# Set figure defaults
knitr::opts_chunk$set(
  fig.width = 6,
  fig.height = 4,
  dpi = 300,
  out.width = "100%"
)

# Set file paths (use relative paths instead of setwd)
data_path <- "/Users/arnavdharmagadda/The Lab Dropbox/Arnav Dharmagadda/gridded_eif_data/"
git_path <- "/Users/arnavdharmagadda/The Lab Dropbox/Arnav Dharmagadda/GitHub/"

# Load packages

if (!requireNamespace("pacman", quietly = TRUE)) {
  install.packages("pacman")
}

pacman::p_load(ncdf4, sf, terra, dplyr, ggplot2, tmap, arrow, dplyr, tidyr, scales, haven, stringr, purrr, tigris, systemfonts, showtext)

# Set default ggplot2 font to Lato (system font)
theme_set(theme_minimal(base_family = "Lato"))

font_add(family = "Lato", regular = "/Users/arnavdharmagadda/Library/Group Containers/UBF8T346G9.Office/FontCache/4/CloudFonts/Lato/25090817022.ttf")  # optional, if not auto-detected
showtext_auto()

```

\leaguegothic\textcolor{custommaroon}{\Huge\textbf{Air Quality: PM2.5 Exposure}}

\vspace{0.5em}

```{=latex}
\setlength{\columnsep}{20pt}
\columnratio{0.60}
\begin{paracol}{2}
```

<!-- LEFT COLUMN -->

\lato\textcolor{black}{\large\textbf{Population Over Space}}

```{r, pm25_map}
file_path <- paste0('/Users/arnavdharmagadda/The Lab Dropbox/Arnav Dharmagadda/GitHub/environmental-inequality-atlas/data/map_data/pm25_with_county.rda')
load(file_path)

# Create hexagonal map with publication styling
hex_map <- tm_basemap("CartoDB.Positron", alpha = 1) +
  tm_shape(hex_grid_sf) +
  tm_fill(
    col = "PM25",
    title = "PM2.5 (μg/m³)",
    style = "cont",
    palette = "-inferno",
    fill_alpha = 0.9,
    colorNA = "transparent",
    textNA = "",
    lwd = 0
  ) +
  tm_borders(
    col = "white",
    lwd = 0.3
  ) +
  tm_graticules(
    n.x = 5,
    n.y = 5,
    col = "gray70",
    lwd = 0.3,
    labels.size = 0.3
  ) +
  tm_layout(
    main.title.size = 0.8,
    main.title.fontfamily = "Lato",
    legend.outside = TRUE,
    legend.text.size = 0.6,
    legend.title.size = 0.7,
    legend.text.fontfamily = "Lato",
    legend.title.fontfamily = "Lato",
    fontfamily = "Lato",
    bg.color = "transparent",
    outer.bg.color = "transparent"
  )

# Display the hexagonal map
print(hex_map)



```

```{=latex}
\switchcolumn
```
```{=latex}
\begin{tcolorbox}[
  colback=white,
  colframe=white,
  boxrule=1pt,
  arc=0pt,
  left=12pt, right=12pt, top=10pt, bottom=10pt
]

\vspace{0em}

\color{black}
\lato{\textcolor{custommaroon}{\large\textbf{Where People Live}}}

\vspace{1em}

\begin{itemize}

\item Charlottesville and Albemarle County are home to 147,428 people. 

\item The most populous county is Albemarle. 

\item The most populous localities are Charlottesville City, Crozet, and Hollymead.

\item The region has a population density of 78 people per square kilometer, which is more dense than the national average of 37 per square kilometer.

\end{itemize}

\vspace{1em}

\end{tcolorbox}
```



```{=latex}
\end{paracol}
```

\vspace{0.5em}

```{=latex}
\setlength{\columnsep}{20pt}
\columnratio{0.55}
\begin{paracol}{2}
```

<!-- LEFT COLUMN -->

\lato\textcolor{custommaroon}{\large\textbf{Population Trends}}

\lato 

* The population has grown since 2000 by 47,506 people at an average growth rate of 1.63%. This is faster than the national average growth rate of 0.8%.


* Albemarle County has the highest growth rate relative to other counties in the region.


```{=latex}
\switchcolumn
```

\lato\textcolor{black}{\large\textbf{Population Over Time}}

```{r, pop_over_time}

file_path <- paste0('/Users/arnavdharmagadda/The Lab Dropbox/Arnav Dharmagadda/GitHub/environmental-inequality-atlas/data/processed/raceincome_rda/raceincome_long.rda')
load(file_path)

raceincome_by_county <- raceincome_combined %>%
  group_by(year, COUNTYFP, NAME) %>%
  summarise(n_noise_postprocessed = sum(n_noise_postprocessed, na.rm = TRUE), .groups = "drop") %>%
  filter(year != 1999)

raceincome_combined <- raceincome_combined %>%
  group_by(year) %>%
  summarise(n_noise_postprocessed = sum(n_noise_postprocessed, na.rm = TRUE)) %>%
  filter(year != 1999)

plot_data <- bind_rows(
  raceincome_by_county,
  raceincome_combined %>% mutate(NAME = "Total", COUNTYFP = "Total")
)

ggplot(plot_data, aes(x = year, y = n_noise_postprocessed, color = NAME)) +
  geom_line(size = 1.2, alpha = 0.9) +
  geom_point(size = 2.5, alpha = 0.9) +
  scale_color_manual(
    values = c(
      "Total" = "#601215",
      "Albemarle" = "#A6CEE3",
      "Charlottesville" = "#003F5C"
    )
  ) +
  labs(
    #title = "Population Over Time",
    x = "Year",
    y = "Population"
  ) +
  scale_x_continuous(
    breaks = seq(2000, 2020, by = 5),
    minor_breaks = NULL,  # Remove minor grid lines
    expand = c(0.02, 0)
  ) +
  scale_y_continuous(
    labels = scales::comma_format(scale = 1e-3, suffix = "K"),
    expand = c(0.02, 0),
    limits = c(0, NA)  # Ensure y-axis starts at zero
  ) +
  theme_minimal(base_size = 12, base_family = "Lato") +
  theme(
    # Plot titles
    plot.title = element_text(
      face = "bold", 
      size = 16, 
      hjust = 0,
      margin = margin(b = 5),
      color = "black"
    ),
    plot.subtitle = element_text(
      size = 14,
      hjust = 0,
      margin = margin(b = 20),
      color = "gray30"
    ),
    plot.caption = element_text(
      size = 9,
      hjust = 0,
      margin = margin(t = 15),
      color = "gray50",
      lineheight = 1.1
    ),
    # Axes
    axis.title = element_text(size = 12, color = "gray20"),
    axis.text = element_text(size = 10, color = "gray30"),
    axis.title.x = element_text(margin = margin(t = 10), size = 12),
    axis.title.y = element_text(margin = margin(r = 10), size = 12),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 10),
    # Grid
    panel.grid.major.x = element_blank(),  # Remove vertical major grid lines
    panel.grid.minor.x = element_blank(),  # Remove vertical minor grid lines
    panel.grid.major = element_line(color = "gray90", size = 0.5),
    panel.grid.minor = element_line(color = "gray95", size = 0.25),
    panel.background = element_rect(fill = "white", color = NA),
    # Add solid x-axis line
    axis.line.x = element_line(color = "black", size = 0.5),
    # Legend
    legend.position = "top",
    legend.title = element_blank(),
    legend.text = element_text(size = 11, color = "gray20"),
    legend.key.width = unit(1.5, "cm"),
    legend.margin = margin(b = 15),
    legend.box.spacing = unit(0, "pt"),
    # Plot margins
    plot.margin = margin(20, 25, 15, 20),
    # Remove plot background
    plot.background = element_rect(fill = "white", color = NA)
  )

```

```{=latex}
\end{paracol}
```

\vspace{0.5em}

```{=latex}
\setlength{\columnsep}{20pt}
\columnratio{0.45}
\begin{paracol}{2}
```

\lato\textcolor{black}{\large\textbf{Age Over Time}}

```{r, age_share}

file_path <- '/Users/arnavdharmagadda/The Lab Dropbox/Arnav Dharmagadda/GitHub/environmental-inequality-atlas/data/processed/ageracesex_rda/ageracesex_year_long.rda'
load(file_path)
data_all <- ageracesex_year_long

data_all <- data_all %>%
  rowwise() %>%
  mutate(total = sum(c(M, F, NA_sex, na_sex), na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(
    black_share = Black / total,
    white_share = White / total,
    hispanic_share = Hispanic / total,
    asian_share = Asian / total,
    male_share = M / total,
    female_share = F / total,
    under_18_share = under_18 / total,
    bet_19_65_share = bet_19_65 / total,
    over_65_share = over_65 / total,
  )

data_collapsed <- data_all %>%
  group_by(year) %>%
  summarise(across(
    c(AIAN, Asian, Black, Hispanic, NA_race, White, other_race, under_18, bet_19_65, over_65, total),
    ~ sum(.x, na.rm = TRUE)
  )) %>%
  mutate(
    under_18_share = under_18 / total,
    bet_19_65_share = bet_19_65 / total,
    over_65_share = over_65 / total
  ) %>%
  filter(year != 1999) %>%
  ungroup()

age_data_long <- data_collapsed %>%
  pivot_longer(cols = c(under_18_share, bet_19_65_share, over_65_share),
               names_to = "type",
               values_to = "value")

ggplot(age_data_long, aes(x = year, y = value, color = type)) +
  geom_line(size = 1.2) +
  geom_point(size = 2.5, alpha = 0.9) +
  scale_color_manual(
    values = c(
      "under_18_share" = "#A6CEE3",
      "bet_19_65_share" = "#601215",
      "over_65_share" = "#003F5C"
    ),
    labels = c(
      "under_18_share" = "Under 18",
      "bet_19_65_share" = "Ages 19-65",
      "over_65_share" = "Over 65"
    )
  ) +
  labs(
    #title = "Age Over Time",
    x = "Year",
    y = "Share of Population"
  ) +
  scale_y_continuous(
    labels = scales::percent_format(),
    expand = c(0.02, 0),
    limits = c(0, 0.8),  # Ensure y-axis starts at zero and goes to 80%
    breaks = seq(0, 0.8, by = 0.2)  # Labels at 0%, 20%, 40%, 60%, 80%
  ) +
  scale_x_continuous(
    breaks = seq(2000, 2020, by = 5),
    minor_breaks = NULL,  # Remove minor grid lines
    expand = c(0.02, 0)
  ) +
  theme_minimal(base_family = "Lato") +
  theme(
    # Axes
    axis.title = element_text(size = 12, color = "gray20"),
    axis.text = element_text(size = 10, color = "gray30"),
    axis.title.x = element_text(margin = margin(t = 10), size = 12),
    axis.title.y = element_text(margin = margin(r = 10), size = 12),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 10),
    legend.position = "top",
    legend.title = element_blank(),
    legend.text = element_text(size = 11, color = "gray20"),
    legend.key.width = unit(1.5, "cm"),
    legend.margin = margin(b = 15),
    legend.box.spacing = unit(0, "pt"),
    plot.title = element_text(size = 16, face = "bold"),
    panel.grid.major.x = element_blank(),  # Remove vertical major grid lines
    panel.grid.minor.x = element_blank(),   # Remove vertical minor grid lines
    # Add solid x-axis line
    axis.line.x = element_line(color = "black", size = 0.5)
  )

```

```{=latex}
\switchcolumn
```
\lato\textcolor{custommaroon}{\large\textbf{Age Dynamics}}

\lato 

* 18% of Charlottesville’s population is under 18, 60% of the population is 19-65, and 20% of the population is over 65.

* Charlottesville has a lower share of people under 18, and a similar share of people between 19 and 65 relative to the US. Charlottesville has a higher share of people over 65 than the US. 

* The over 65 population in Charlottesville is growing, while the under 18 and 19-65 population is declining.

```{=latex}
\end{paracol}
```